---
title: "p8108_hw6_vas2145"
author: "Victoria Schliep"
date: "2023-12-02"
output: github_document
---


#loading libraries
```{r}
library(tidyverse)
library(modelr)
library(mgcv)
library(dplyr)
library(knitr)
set.seed(1)
```

#problem 2
#download provided data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

#create 5,000 bootstrap samples
```{r}
bootstrap_df = weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    rsq = map(models, broom::glance)) |> 
  select(results, .id, rsq) |> 
  unnest(results) |> 
  filter(term %in% c("tmin", "prcp")) |> 
  group_by(.id) |> 
  mutate(beta1xbeta2 = prod(estimate),
         log_beta_product = log(beta1xbeta2)) |> 
  select(log_beta_product, rsq) |> 
  unnest(rsq) |> 
  janitor::clean_names() |> 
  select(log_beta_product, id, r_squared) |> 
  unique()

```

#Plot the distribution of estimates

```{r}
bootstrap_df |> 
  ggplot(aes(x = log_beta_product)) +
  geom_density()
```

The plot of the distribution estimates is a single peak (unimodal) plot with a peak at approximately -6.25 log of the beta product and at 0.4 density. It is a left skewed distribution.

```{r}
bootstrap_df |> 
  ggplot(aes(x = r_squared)) + geom_density()
```

The plot depicts the distribution estimates of the r-squared value. It is a unimodal plot with a peak at approximately 0.90 r_squared and 28 density.

```{r}
bootstrap_ci = bootstrap_df |> 
  unique() |>
  ungroup() |> 
  select(-id) |> 
  summarize(beta_mean = mean(log_beta_product, na.rm = TRUE),
            beta_high = quantile(log_beta_product, 0.975, na.rm = TRUE),
            beta_low = quantile(log_beta_product, 0.025, na.rm = TRUE),
            rs_mean = mean(r_squared),
            rs_high = quantile(r_squared, 0.975),
            rs_low = quantile(r_squared, 0.025))

bootstrap_ci |> knitr::kable()

```
The 95% confidence interval for the log of the beta product is (`r bootstrap_ci$beta_low`, `r bootstrap_ci$beta_high`).

The 95% confidence interval for r-squared is (`r bootstrap_ci$rs_low`, `r bootstrap_ci$rs_high`).

#Problem 3

Download data

```{r}
bwt_df = read.csv("birthweight.csv")

bwt_df = bwt_df |> 
  janitor::clean_names() |> 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)) |> 
  drop_na()

```

Create a regression model

```{r}
fit = lm(bwt ~ fincome + gaweeks + smoken, data = bwt_df)

fit |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  knitr::kable(digits=3)

summary(fit)

```

The regression model is based on key factors that I believe may have an influence on a child's birthweight, the family's monthly income (`fincome`), the gestational weeks in age0  (`gaweeks`) and the average number of cigarettes smoked per day in pregnancy (`smoken`). Because the p-values are all less than 0.01, none need to be removed from the model.

```{r}
bwt_df |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  ggplot(aes(x = resid, y = pred)) +
  geom_point(alpha = 0.5)
```

